% Abstract. N=150 stratified + N=30 full abstention + N=21 SSU + efficiency.
Long-horizon conversational agents need a memory that is not only
\emph{searchable} but also \emph{truthful}: when the relevant fact was
never stated, the agent should abstain; when a preference was updated,
the agent should retrieve the current value, not the stale one. We
revisit agentic memory through the lens of \emph{truth maintenance}
and propose the \textbf{Temporal Truth-Maintained Memory Graph}
(TTMG), a claim-level memory that augments a self-organising note
graph with validity intervals, provenance, polarity, and explicit
\emph{contradict} and \emph{supersede} edges. A lightweight LLM
linker gated by a subject--predicate lexical check writes these
edges; an intent-gated retriever returns a truth-consistent subgraph
of active claims, with a compute-free raw-turn fallback for
non-temporal questions. We evaluate TTMG against a flat hybrid-RAG
baseline and our A-Mem reimplementation on LongMemEval-S
(DeepSeek-V3.2 reader, MiniLM-L6 embeddings, identical retrieval
budget).  On the \textbf{single-session-user} capability slice
($n{=}21$), TTMG answers 100\% of questions correctly (paired 2--0
against Flat RAG's \FlatLMEAccSSU\%), a strict set-inclusion
improvement bought directly by the claim schema's subject/
\texttt{valid\_from} framing. Ablating only the LLM linker drops
\emph{every} measured slice: overall falls from \TTMGLMEAcc\% to
\AblContradict\% ($-$2.6~pp), abstention by $-$22~pp,
knowledge-update by $-$9~pp, and single-session-user by $-$5~pp,
confirming the linker's causal role. On the full 30-question abstention set, TTMG
\emph{explicitly abstains} on \TTMGABSFullAbstainRate\% of
questions while Flat RAG never does, delivering the same outcome
accuracy (\TTMGABSFull\% each) via a behaviourally different
policy. Overall accuracy on the 150-question stratified slice is
statistically indistinguishable from Flat RAG's (paired McNemar
$p{=}$\McnemarOverall); composing TTMG with Flat RAG via the
inference-time TTMG-Abstain router (\AbstainRouterAcc\%) or the
Kimi-K2 zero-shot-intent TTMG+R-LLM router (\LLMRouterAcc\%) places
TTMG's truth-maintenance behaviours into a system that matches Flat
RAG overall. TTMG runs at \TTMGTokens{} tokens per question versus
\AmemTokens{} for A-Mem (\TokensOverhead\% relative), so the added
semantics come at strictly lower cost.
