\section{Related Work}
\label{sec:related}

\paragraph{Agentic memory systems.}
Early LLM agent memories treat conversation history as a flat vector
index \citep{mem0,longrag}. MemGPT~\citep{memgpt} introduces an
OS-style hierarchical memory controller that pages between a small
fast context and a larger recall store; its mechanism is
\emph{retrieval-time} eviction, with no explicit model of
contradicting or superseded content. MemoryBank~\citep{memorybank}
introduces an Ebbinghaus-curve forgetting mechanism with periodic
consolidation of frequently-recalled notes; the consolidation step
is a summary rewrite, not a truth-maintenance operation. Both
treat the note/chunk as the atomic unit and rely on downstream
retrieval to pick the currently relevant entry. A-Mem~\citep{amem}
goes further by treating memory as a self-organising graph of
structured notes, each carrying keywords, contextual descriptions,
and tags, with links that are formed and rewritten as new notes
arrive. These systems are \emph{retrieval-centric}: they surface
the closest matching textual chunk. They have no explicit handle on
\emph{when} a statement was true, \emph{whether} a later statement
replaced it, or \emph{why} two statements might be incompatible.
TTMG shares A-Mem's skeleton -- LLM-enriched nodes with
automatically discovered links -- but changes the unit of memory
from a \emph{note} to a \emph{claim} and the class of links from
semantic similarity to \emph{truth-maintenance relations}
(\texttt{support}, \texttt{contradict}, \texttt{supersede}).
Memory consolidation \citep{memoria} and hierarchical memory
\citep{hippocampal} add compression on top and are orthogonal to
the truth-maintenance mechanism.

\paragraph{Truth maintenance and belief revision.}
Classical TMSs~\citep{tms_doyle,atms_dekleer} and belief-revision
calculi~\citep{agm,nonmonotonic} maintain the consistency of a
logical knowledge base as new assertions arrive. TTMG imports the
core idea -- every assertion carries a validity interval and
provenance, and the graph tracks which assertion is currently in
force -- but uses an LLM-based pairwise labeler instead of a formal
entailment solver, so it runs inside a dialogue system.

\paragraph{Temporal knowledge graphs and contradiction detection.}
Temporal knowledge graphs \citep{tkg_survey,tkgqa} attach time stamps
to $(s, p, o)$ edges; contradiction detection has a long NLI-based
line \citep{nli_contradict,anli}. TTMG differs in being a \emph{live}
agent memory: claims are generated and linked as the agent talks,
the contradiction scorer is a small LLM call whose output directly
flips a claim's \texttt{active} bit, and the retriever computes a
maximum-consistent subgraph over returned candidates.

\paragraph{Long-context conversational benchmarks.}
LongMemEval~\citep{longmemeval} contributes 500 questions across six
capabilities with session-level answer attribution and 30 abstention
probes. LoCoMo~\citep{locomo} extends dialogue length and
MemoryAgentBench~\citep{memoryagentbench} targets conflict
resolution; we focus on LongMemEval-S because its abstention and
knowledge-update slices isolate the failure modes TTMG targets.
