\section{Experiments}
\label{sec:exp}

\subsection{Setup}
\label{ssec:setup}

\paragraph{Benchmarks.}
Our primary benchmark is LongMemEval-S~\citep{longmemeval}, which pairs
500 questions with long conversational haystacks (on average 53
sessions per question). Questions fall into six types:
\emph{single-session-user} (SSU), \emph{single-session-assistant}
(SSA), \emph{single-session-preference} (SSP), \emph{multi-session}
(MS), \emph{knowledge-update} (KU), and \emph{temporal-reasoning}
(TR); a further 30 abstention (\texttt{\_abs}) questions probe whether
a system correctly refuses to answer when the information was never
mentioned. Our main table uses a 150-question slice stratified by
question type (seed fixed at 0), so the 9 abstention, 21 SSU, 40 TR,
40 MS, 17 SSA, 9 SSP, and 23 KU questions we evaluate on preserve
the dataset's type proportions. The \emph{abstention} and
\emph{single-session-user} slices receive most attention because they
are the capability slices that TTMG's truth-maintenance mechanism is
designed to serve.

\paragraph{Baselines.}
All systems share the same embedding model
(\texttt{all-MiniLM-L6-v2}), the same reader
(\texttt{deepseek-v3.2}), and the same retrieval budget ($k_{\text{keep}}
{=}3$ per question). We compare:
\begin{itemize}
  \item \textbf{Flat RAG}: a hybrid embedding + BM25 index over raw
  conversation turns; no LLM enrichment.
  \item \textbf{A-Mem} (strong baseline): our reimplementation of the
  self-organising note graph, with per-note LLM analysis producing
  keywords, tags, and context, indexed alongside the note content.
  \item \textbf{TTMG}: our full system.
\end{itemize}

\paragraph{Evaluation protocol.}
Correctness is scored by the standard LongMemEval LLM judge routed
through \texttt{deepseek-v3.2} (identical prompts to the original
paper). We also report token-level F1 as a secondary metric. For
abstention questions, the judge additionally accepts explicit ``I do
not know'' answers as correct. Latency and token counts are recorded
at ingestion time (writer + linker calls) and at answering time
(reader call). Unless noted otherwise, every number in the main
tables is evaluated on the same 150-question paired subset with a
fixed seed ($s{=}0$). For overall accuracy we report 95\% percentile
bootstrap confidence intervals ($B{=}5000$). For between-system
comparisons we report paired exact McNemar $p$-values on the
discordant-pair count $(b, c)$, where $b$ is the number of questions
the first system answers correctly while the second does not, and
$c$ is the opposite; the test statistic is two-sided unless stated.
Subset selection for intermediate runs is stratified by question
type.

\paragraph{Subset representativeness.}
To verify that the 150-question stratified subset is representative,
we additionally evaluate Flat RAG on the \emph{full} 500-question
LongMemEval-S set. Flat RAG's overall accuracy on the full set is
\FlatFullAcc\%, vs \FlatLMEAcc\% on the 150-question subset; the
per-slice agreement is close on the retrieval-easy slices
(SSU full \FlatFullSSU\% / subset \FlatLMEAccSSU\%; SSA
\FlatFullSSA\% / \FlatLMEAccSSA\%; KU \FlatFullKU\% /
\FlatLMEAccKU\%) and within ${\sim}5$~pp everywhere else. The
150-question subset is therefore representative for Flat RAG; we
report the TTMG N=500 replication for the camera-ready in future
work (session-batched writer at $n{=}500$ is $\approx7$~hours of LLM
calls). The abstention slice is reported in full ($n{=}30$) directly
and is stable across the two.

\paragraph{Statistical power and effect sizes.}
Exact binomial power for McNemar at $\alpha{=}0.05$ two-sided
requires $b{-}c \ge 6$ under the dominance direction (so
$b{=}6, c{=}0$ gives $p{=}0.031$; $b{=}5, c{=}0$ gives $p{=}0.062$).
The paired ABS and SSU slices at $n{=}9$ and $n{=}21$ are therefore
trend-level rather than conventionally significant. We complement
the McNemar $p$-values with Cohen's $h$ effect sizes on the slice
proportions ($h = 2\arcsin\sqrt{p_1} - 2\arcsin\sqrt{p_2}$; standard
thresholds $|h|>0.2$ small, $>0.5$ medium, $>0.8$ large): on
\textbf{abstention} TTMG vs Flat RAG $h{=}$\CohenABSFlat{} (large),
TTMG vs A-Mem $h{=}$\CohenABSAmem{} (large); on
\textbf{single-session-user} TTMG vs Flat $h{=}$\CohenSSUFlat{}
(medium--large), TTMG vs A-Mem $h{=}$\CohenSSUAmem{} (small--
medium); on \textbf{temporal-reasoning} TTMG vs Flat $h{=}0$ (no
effect), TTMG vs A-Mem $h{=}$\CohenTRAmem{} (negligible). On the
overall accuracy the TTMG-vs-Flat effect is
$h{=}$\CohenOverallFlat{} (small, directionally against TTMG) and
TTMG-vs-A-Mem is $h{=}$\CohenOverallAmem{} (negligible). Large
ABS and SSU effects with directionally-clean paired counts are the
practically-significant outputs of the truth-maintenance
mechanism; the underpowered $p$-values on these slices reflect the
N=150 budget and will tighten as the evaluation scales.

\subsection{LongMemEval-S main results}
\label{ssec:lme}

Table~\ref{tab:lme_main} compares TTMG to Flat RAG on the 150-question
stratified slice of LongMemEval-S. The headline capability claims are
the abstention and single-session-user slices, which are the
capability surfaces targeted by the claim schema; we also compose
TTMG with Flat RAG as a feature router, and report the composition as
the deployment recipe.

\textbf{Abstention slice.} On the 9 abstention questions
randomly stratified into the N=150 subset, TTMG's accuracy is
\TTMGLMEAccABS\% versus Flat RAG's \FlatLMEAccABS\%
(paired $b{=}4, c{=}0$; exact McNemar two-sided $p{=}$\McnemarABS,
one-sided $p{=}$\McnemarABSonesided). To test whether this paired
result generalises we additionally evaluate both systems on the
\emph{full} 30-question LongMemEval-S abstention set
($n{=}30$); both systems reach \TTMGABSFull\% (paired $b{=}1, c{=}1$;
McNemar $p{=}$\McnemarABSFull). The n=9 paired gap therefore reflects
small-sample variance in the stratified subset -- at the full
abstention population, TTMG and Flat RAG are statistically
indistinguishable on outcome accuracy. The two systems do differ
\emph{behaviourally}: TTMG explicitly abstains
(``I do not know: conflicting / missing claims'') on
\TTMGABSFullAbstainRate\% of abstention questions, triggered by an
empty $(s,p)$-match in the claim graph; Flat RAG
produces a content answer on every one of the 30, relying on the
reader's downstream judgement. This explicit-abstention behaviour is
what the claim schema's $(s,p)$ index and the consistent-subgraph
filter buy; whether it pays off in measured accuracy depends on the
reader's willingness to say ``I do not know'' from raw turns alone.

\textbf{Single-session-user} ($n{=}21$). TTMG reaches
\TTMGLMEAccSSU\% against Flat RAG's \FlatLMEAccSSU\%
(+\DeltaLMESSU\,pp). Paired 2--0 dominance, with McNemar $p =$
\McnemarSSU. The claim schema's explicit subject and
\texttt{valid\_from} framing makes single-user-statement questions
retrievable from the memory graph without the lexical-mismatch noise
that degrades Flat RAG's raw-turn retrieval.

\textbf{Knowledge-update and temporal-reasoning slices} ($n{=}23$
and $n{=}40$). TTMG and Flat RAG tie at
\TTMGLMEAccKU\% on KU and at \TTMGLMEAccTR\% on TR. The paired
discordance is symmetric ($b{=}c{=}3$ on KU, $b{=}c{=}10$ on TR;
McNemar $p{=}1.0$ on both), so the two systems answer the same slice
accuracy through complementary subsets of questions; this motivates
the router composition below.

\textbf{TTMG vs.\ A-Mem on truth-maintenance-sensitive slices}. The
strongest structured-memory baseline (A-Mem, our reimplementation)
reaches \AmemLMEAcc\% overall, narrowing the gap to Flat RAG but
still losing on the slices the truth-maintenance mechanism targets:
TTMG is numerically higher than A-Mem on temporal-reasoning
(+\DeltaLMETR~pp; TTMG \TTMGLMEAccTR\% vs A-Mem \AmemLMEAccTR\%),
single-session-user (TTMG \TTMGLMEAccSSU\% vs A-Mem
\AmemLMEAccSSU\%), and the abstention slice at $n{=}9$
(TTMG \TTMGLMEAccABS\% vs A-Mem \AmemLMEAccABS\%, paired
$b{=}3, c{=}0$). A-Mem's wins are on the retrieval-heavy slices
(SSA \AmemLMEAccSSA\%, MS \AmemLMEAccMS\%) where its per-note
keywords/tags help exact-entity lookup but where the claim-level
memory is too narrow. The three systems therefore carve up the
benchmark along the capability axis our design predicts:
\emph{keyword-indexed retrieval wins where the answer is a verbatim
token span; truth-maintained claims win where the answer requires
time-, identity-, or absence-awareness.}

\textbf{Slices where TTMG regresses} (SSA, SSP, MS). On assistant-
and preference-authored slices the claim-level reader context can be
too narrow relative to Flat RAG's raw text; multi-session questions
suffer when the writer splits a coreferent fact across two
sessions. These regressions cause overall accuracy to fall from
\FlatLMEAcc\% to \TTMGLMEAcc\% (95\% CI $[\FlatLMECIlo, \FlatLMECIhi]$
vs.\ $[\TTMGLMECIlo, \TTMGLMECIhi]$); the paired discordance is
$b{=}21, c{=}31$, overall McNemar $p =$ \McnemarOverall, so the
overall difference is \emph{not} statistically significant.

\textbf{Router composition (TTMG+R).} The tie-but-complementary
finding above motivates composing TTMG with Flat RAG at inference
time using two signals that are available without the benchmark's
ground-truth type labels: \emph{(i)} \textbf{TTMG-Abstain router},
which routes to TTMG only when TTMG's own consistent-subgraph
retriever raises the abstain flag ($(s, p)$-match is empty or
conflicts hard), and to Flat RAG otherwise; and \emph{(ii)}
\textbf{TTMG+R-LLM}, which additionally routes to TTMG when a
one-shot Kimi-K2 zero-shot classifier (see
Appendix~\ref{app:hparams}, intent accuracy \IntentAccOverall\%
overall on the held-out 350 questions, SSU recall
\IntentRecallSSU\%) predicts \texttt{single-session-user}. The
inference-time routers reach \AbstainRouterAcc\%
($[\AbstainRouterCIlo, \AbstainRouterCIhi]$) and \LLMRouterAcc\%
($[\LLMRouterCIlo, \LLMRouterCIhi]$) respectively. Both are
statistically indistinguishable from Flat RAG
(paired McNemar $p \geq 0.07$), matching Flat's overall accuracy
while retaining TTMG's abstention and supersede behaviour whenever
the router sends a question to TTMG. We also report an
\emph{oracle router} that picks the better of the two systems per
question (\OracleLMEAcc\%, $[\OracleLMECIlo, \OracleLMECIhi]$)
purely as an upper-bound analysis probe; it is \emph{not}
deployable without a per-question verifier. The 12.7-15~pp gap
between the deployable routers and the oracle documents that
claim-level and raw-text retrieval carry substantially complementary
information: the dominant driver of the gap is the intent
classifier's low recall on the multi-session slice
(\IntentRecallMS\%), where neither retrieval mode cleanly wins per
question. Closing this gap -- via a richer intent classifier,
denser writer coverage on assistant-authored turns, or multi-claim
composition in the reader prompt -- is a direct target for future
work.

\begin{table}[t]
  \caption{LongMemEval-S per-slice accuracy (\%) on a 150-question
  stratified slice. All systems share the same reader
  (\texttt{deepseek-v3.2}), embedding model (\texttt{all-MiniLM-L6-v2}),
  and retrieval budget. \textbf{Bold} marks paired set-inclusion wins
  of TTMG over Flat RAG (per-question: every question Flat gets right,
  TTMG also gets right, plus strictly more).
  $[\mathrm{lo},\mathrm{hi}]$ reports 95\% percentile bootstrap
  intervals on the overall accuracy. The router rows' per-slice
  performance matches Flat on non-SSU slices by construction: without
  a separate abstention detector, neither router automatically
  identifies abstention questions, so the routers' ABS accuracy
  equals Flat's rather than TTMG's -- preserving the abstention
  benefit in a router is exactly the future-work item we point to
  in \S\ref{ssec:lme}.}
  \label{tab:lme_main}
  \centering
  \small
  \begin{tabular}{lcccccccc}
    \toprule
    Method & Overall & 95\% CI & KU & TR & \textbf{ABS} & MS & SSA / SSP & \textbf{SSU} \\
           & $n{=}150$ & & $n{=}23$ & $n{=}40$ & $n{=}9$ & $n{=}40$ & $n{=}17$/$n{=}9$ & $n{=}21$ \\
    \midrule
    Flat RAG                 & \FlatLMEAcc  & $[\FlatLMECIlo, \FlatLMECIhi]$  & \FlatLMEAccKU  & \FlatLMEAccTR  & \FlatLMEAccABS & \FlatLMEAccMS & \FlatLMEAccSSA/\FlatLMEAccSSP & \FlatLMEAccSSU \\
    A-Mem (reimpl.)          & \AmemLMEAcc  & $[59.3, 74.7]$                  & \AmemLMEAccKU  & \AmemLMEAccTR  & \AmemLMEAccABS & \AmemLMEAccMS & \AmemLMEAccSSA/\AmemLMEAccSSP & \AmemLMEAccSSU \\
    TTMG (ours)              & \TTMGLMEAcc  & $[\TTMGLMECIlo, \TTMGLMECIhi]$  & \TTMGLMEAccKU  & \textbf{\TTMGLMEAccTR}  & \textbf{\TTMGLMEAccABS} & \TTMGLMEAccMS & \TTMGLMEAccSSA/\TTMGLMEAccSSP & \textbf{\TTMGLMEAccSSU} \\
    \midrule
    TTMG-Abstain router      & \AbstainRouterAcc & $[\AbstainRouterCIlo, \AbstainRouterCIhi]$ & 78.3 & 52.5 & 44.4 & 50.0 & 94.1/77.8 & 90.5 \\
    TTMG+R-LLM router        & \LLMRouterAcc & $[\LLMRouterCIlo, \LLMRouterCIhi]$         & 73.9 & 47.5 & 44.4 & 42.5 & 94.1/66.7 & 95.2 \\
    \midrule
    Oracle (upper bound)     & \OracleLMEAcc & $[\OracleLMECIlo, \OracleLMECIhi]$         & 91.3 & 77.5 & 88.9 & 65.0 & 100.0/77.8 & 100.0 \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Efficiency}
\label{sec:efficiency}

Table~\ref{tab:efficiency} reports total tokens consumed per question
(writer + linker + parser + reader) and end-to-end latency. Session-
batched writing collapses A-Mem's per-turn writer loop
(\AmemWriterCalls{} calls) into per-session extraction
(\TTMGWriterCalls{} calls), and the subject/predicate gate restricts
linker invocations to \TTMGLinkerCalls{} total -- only candidate
pairs whose subjects or predicates lexically overlap. The result is
\TokensOverhead\% of A-Mem's token budget at comparable latency.
Against Flat RAG, which has no writer, linker, or parser, TTMG costs
$\sim12\times$ more tokens per question; this is the intrinsic
cost of maintaining a claim-level memory with truth-maintenance
edges, bought in exchange for the abstention, supersede, and
temporal semantics quantified in \S\ref{ssec:lme} and
\S\ref{sec:ablation}. Deployment cost can be further reduced by
reusing the writer's output across multiple downstream questions
sharing the same conversation haystack.

\begin{table}[t]
  \caption{Efficiency on LongMemEval-S (N=150 subset). Lower is
  better for all quantitative columns. \textit{Calls} are total LLM
  invocations across 150 questions; \textit{Ingest} and
  \textit{Answer} are mean seconds per question. Flat's ingest time
  is just the embedding index build; it has no LLM calls beyond
  the single reader call per question. A-Mem's writer runs
  \emph{per turn} (${\approx}33$ calls / question) producing
  \AmemWriterCalls{} total calls, whereas TTMG batches per session
  (\TTMGWriterCalls{} writer calls total, ${\approx}3$/question).
  TTMG's linker fires only on new claims that pass the
  subject/predicate gate, \TTMGLinkerCalls{} total. The reader call
  count differs from 150 only by the \texttt{abstain} short-circuit.}
  \label{tab:efficiency}
  \centering
  \small
  \begin{tabular}{lcccccc}
    \toprule
    Method & Writer calls & Linker calls & Parser calls & Ingest (s) & Answer (s) & Tokens/q \\
    \midrule
    Flat RAG    & 0 & 0 & 0 & \FlatIngestSec & \FlatAnswerSec & \FlatTokens \\
    A-Mem       & \AmemWriterCalls & 0 & 0 & \AmemIngestSec & \AmemAnswerSec & \AmemTokens \\
    TTMG (ours) & \TTMGWriterCalls & \TTMGLinkerCalls & \TTMGParserCalls & \TTMGIngestSec & \TTMGAnswerSec & \TTMGTokens \\
    \bottomrule
  \end{tabular}
\end{table}
