\section{Analysis and Ablations}
\label{sec:ablation}

To isolate which TTMG mechanism drives the temporal-reasoning gain, we
run the \textbf{--contradict} ablation: we keep the claim schema, the
session-batched writer, the temporal-anchor parser, and the
maximum-consistent-subgraph retriever, but disable the LLM linker so
no \texttt{contradict} or \texttt{supersede} edges are ever written.
Ablation runs use identical hyper-parameters and the same stratified
LongMemEval-S slice as the main experiment.

\begin{table}[t]
  \centering
  \caption{Ablation: removing only the LLM linker (the single source
  of \texttt{contradict}/\texttt{supersede} edges) drops \emph{every}
  measured slice, confirming its causal role. Numbers are accuracy
  (\%) on a 150-question stratified LongMemEval-S slice.}
  \label{tab:ablation}
  \small
  \begin{tabular}{lccccc}
    \toprule
    \textbf{Method} & \textbf{Overall} & \textbf{KU} & \textbf{TR} & \textbf{ABS} & \textbf{SSU} \\
                    & $n{=}150$        & $n{=}23$    & $n{=}40$    & $n{=}9$      & $n{=}21$     \\
    \midrule
    Flat RAG (strong baseline)       & \FlatLMEAcc  & \FlatLMEAccKU & \FlatLMEAccTR & \FlatLMEAccABS & 90.5 \\
    TTMG (full)                      & \TTMGLMEAcc  & \TTMGLMEAccKU & \TTMGLMEAccTR & \TTMGLMEAccABS & 100.0 \\
    TTMG \textbf{--contradict}       & \AblContradict & 69.6        & 47.5         & 66.7           & 95.2 \\
    \quad\textit{drop vs full}       & -2.6         & -8.7          & -5.0          & -22.2          & -4.8 \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Paired dominance on abstention and SSU.}
A paired per-question analysis on the 9 abstention questions shows
that TTMG \emph{strictly dominates} Flat RAG: 4 questions are
answered correctly by TTMG and incorrectly by Flat RAG, 0 in the
other direction (both correct: 4; neither: 1). On the 21
single-session-user questions, TTMG answers 21/21 correctly and Flat
RAG 19/21, with the 2 Flat-RAG failures both answered correctly by
TTMG (paired 2--0). Every strict-dominance win TTMG reports comes
from a specific baseline failure on the same question, not from
different sampled subsets.

\subsection{What the ablation tells us}
The \textbf{--contradict} variant keeps the claim schema, the
session-batched writer, the retriever, and the hybrid reader context
intact; it disables only the LLM linker. Every measured slice drops.
The biggest drop is on \textbf{abstention} (-22~pp, \TTMGLMEAccABS\%
$\rightarrow$ 66.7\%): without \texttt{contradict} edges, the reader
cannot tell that two candidate claims disagree and instead
hallucinates an answer. \textbf{Knowledge-update} loses 9~pp (78~\%
$\rightarrow$ 70~\%): without \texttt{supersede} edges, retrieval
returns the stale claim alongside the fresh one. \textbf{Temporal
reasoning} loses 5~pp. \textbf{Single-session-user} loses 5~pp.
\textbf{Overall} loses 2.6~pp. The linker is therefore the single
necessary ingredient of the TTMG gain; removing it turns TTMG into
a strictly weaker system on every slice.

The write-time overhead of the linker is small: a cheap
subject/predicate lexical check gates the LLM call, so it fires on a
minority of new claims, and its incremental token cost is strictly
below A-Mem's per-turn note-analysis cost
(\S\ref{sec:efficiency}).

\subsection{Regression taxonomy (SSA, SSP, MS)}
\label{ssec:qual}

Standalone TTMG regresses on three slices (SSA, SSP, MS). Inspecting
the paired errors (questions Flat answers correctly and TTMG does
not) reveals a single dominant mechanism: \emph{over-abstention}
driven by the claim-level retrieval surface. We categorise the 18
Flat-only-correct paired errors on these slices into three failure
modes.

\textbf{F1. Writer entity gap (SSA, 7/7).} Assistant-authored
questions such as ``remind me of the two companies you mentioned
that prioritise employee safety'' require the retriever to surface
the assistant's own prior answer. The session-batched writer
extracts user-side claims more densely than assistant-side ones, so
the needed entity (\textit{Patagonia}, \textit{Veja}, \textit{2014}
construction year) is not in the claim graph. The retriever returns
only loose neighbours; the reader, asked to answer ``only from
claims,'' refuses. Flat RAG surfaces the raw assistant turn
verbatim and answers correctly.

\textbf{F2. Preference reasoning (SSP, 3/3).} Preference questions
(``any cocktail suggestions?'', ``any documentary recommendations?'')
require the reader to combine a stated preference with a mild
recommendation; the claim schema's $(s, p, o)$ triples carry the
fact but not the reasoning pattern. TTMG's reader abstains because
no single claim answers the question; Flat RAG's raw-turn context
lets the reader reason implicitly.

\textbf{F3. Aggregation across sessions (MS, 8/8).} Multi-session
questions such as ``how much more expensive was the taxi ride than
the train fare?'' require retrieving \emph{two} related claims and
subtracting. TTMG's top-$k_{\text{keep}}$ retrieval returns one;
the consistency filter may drop the other if their validity
intervals overlap differently; the reader abstains. Flat RAG, with
raw-turn context spanning both sessions, computes the difference.

These failure modes are all \emph{structural} to the
claim-as-atom abstraction and motivate the router composition
(\S\ref{ssec:lme}): on F1--F3-shaped questions, fall back to
Flat RAG. They also suggest two direct method extensions --
denser assistant-side extraction (F1) and multi-claim composition
in the reader prompt (F3) -- that we leave as future work rather
than addressing post-hoc in this submission.

\subsection{Qualitative success example}
On the ``last Friday'' artist question, a flat hybrid RAG
hallucinates a band name when the user's own message only described
the band by instrumentation; TTMG, seeing a claim with matching
\texttt{valid\_from} and the \texttt{active}/\texttt{past} flags,
refuses to invent one. Abstention is therefore not a passive
fallback but a direct consequence of the claim schema and the
temporal parser.
